---
title: "Eurostat"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{eurostat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
library(pddcs)
indicators <-
  indicatorlist[indicatorlist$source == 'eurostat',]$indicator
```

This vignette provides details on the source data retrieved from Eurostat.

```{r, eval=TRUE, message=FALSE}
library(pddcs)
```

### About the source

`fetch_indicator()` with `source = 'eurostat'` uses the R package [eurostat](https://ropengov.github.io/eurostat/index.html) to retrieve data.

Data is retrieved with `eurostat::get_eurostat()` with `keepFlags = TRUE` to ensure that footnotes are included in the returned data. 

Note that values are only returned for EU countries (excluding Cyprus), as well as UK, Norway and Island.

### Available indicators

Currently only the following two indicators are available for Eurostat. Both indicators are from the `demo_gind` dataset in the [Eurostat list of databases](https://ec.europa.eu/eurostat/web/main/data/database).

```{r, echo=TRUE, message=FALSE}
library(dplyr)
data(indicatorlist)
indicatorlist %>% 
  filter(source == 'eurostat') %>% 
  select(indicator)
```

### Examples

```{r, echo=TRUE, message=FALSE, eval=FALSE}
# Retrieve population and birth rate data
df <- fetch_indicator(
  indicator = c('SP.POP.TOTL', 'SP.DYN.CBRT.IN'),
  source = 'eurostat')
```

### References

Leo Lahti, Przemyslaw Biecek, Markus Kainu and Janne Huovari. Retrieval and analysis of Eurostat open data with the eurostat package. R Journal 9(1):385-392, 2017. R package version 3.7.1. URL: <https://ropengov.github.io/eurostat/>
