---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette gives an introduction to the `pddcs` package.

```{r setup}
library(pddcs)
```

## Fetch data

The most important function in `pddcs` is `fetch_indicator()`.

`fetch_indicator()` retrieves data for all available sources and indicators. It takes two basic arguments; `indicator`, a [CETS](https://datahelpdesk.worldbank.org/knowledgebase/articles/201175-how-does-the-world-bank-code-its-indicators) type indicator code, and `source`, a specified data source. Currently available sources are [Eurostat](https://ec.europa.eu/eurostat/data/database), [UNPD WPP](https://population.un.org/wpp/), [UNICEF](https://data.unicef.org/) and [WHO](https://www.who.int/data/gho). For a list of available indicators see the package dataset `indicatorlist`.

```{r, eval=FALSE}
# Check for available indicators 
?indicatorlist
```

In order to fetch the most recent population data from Eurostat, we simply specify the indicator code (`'SP.POP.TOTL'`) and the source (`'eurostat'`).

```{r}
# Fetch population data from Eurostat
df <- fetch_indicator('SP.POP.TOTL', source = 'eurostat')
```

`fetch_indicator()` always returns a data frame with `r ncol(df)` columns; `iso3c` (country code) , `year` (year) , `indicator` (indicator code), `value` (data value), `note` (footnote) and `source` (data source).

```{r}
# Inspect data frame 
head(df)
```

## Compare with WDI

Before uploading any data to The Data Collection System (DCS) it can be useful to compare it with the current data in WDI. This can be done with `compare_with_wdi()`, which only takes one argument; a data frame as returned by `fetch_indicator().`

```{r}
# Compare with WDI
dl <- compare_with_wdi(df)
```

`compare_with_wdi()` returns a list of three data frames; the original dataset (`source`), the data retrieved from WDI (`wdi`) and the rows in the source dataset that are not present in WDI (`not_in_wdi`).

```{r}
# Inspect list
str(dl)
```

Here we select the rows that don't match.

```{r}
# Select rows with 'new' data
df <- dl$not_in_wdi
```

## Format data

DCS requires a very specific format in order to upload files to the system. `pddcs` has a set of functions to help to help simplify this process. `format_dcs()` takes two arguments; `df` (a data frame in `pddcs` format) and `type`, which specifies if the the dataset should be convert to data or metadata format.

```{r}
# Convert to DCS 'data' format 
df <- format_dcs(df, type = 'data')
head(df)
```

## Write data

We can then write this data frame to file of our choosing with `write_dcs()`.

`write_dcs()` checks that the data has the correct format and ensure that the resulting file has the correct sheet names (i.e. `'Sheet1'` for data and `'Country-Series-Time_Table'` for metadata).

```{r, eval=FALSE}
# Write to a DCS formatted file  
write_dcs(df, path = 'data-SP.POP.TOTL-eurostat.xlsx', type = 'data')
File saved to data-SP.POP.TOTL-eurostat.xlsx.
```


