---
title: "UN Population Division"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{unpd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(echo = TRUE)
```

This vignette provides details on the source data retrieved from UNPD.

```{r, eval=TRUE, message=FALSE}
library(pddcs)
```

### About the source

`fetch_indicator()` with `source = 'who'` retrieves data from a series of online Excel sheets from [World Population Prospects 2019](https://population.un.org/wpp/) provided by UN's Population Division.

Both Estimate and Medium variants are retrieved. Data values for the years 2020 - 2050 are estimates.

The use of Excel sheets as source data is likely to change when a new UN API is released.

### Available indicators

Currently the following indicators are available for UNPD.

```{r, echo=TRUE, message=FALSE}
library(dplyr)
data(indicatorlist)
indicatorlist %>% 
  filter(source == 'unpd') %>% 
  select(indicator)
```

Note that the code `SP.POP.AG00.MA.IN` is an exception to how `fetch_indicator()` usually works. Using `indicator = 'SP.POP.AG00.MA.IN'` will return data for all population groups from 0 - 25 years (i.e SP.POP.AG00.MA.IN - SP.POP.AG25.MA.IN).

### Examples

```{r, echo=TRUE, message=FALSE, eval=FALSE}
# Fetch population data from UNPD
df <- fetch_indicator('SP.POP.TOTL', source = 'unpd')

# Fetch male population data for all age groups  (0 - 25)
df <- fetch_indicator('SP.POP.AG00.MA.IN', source = 'unpd')
```

### References

UN. World Population Prospects 2019. <https://population.un.org/wpp/>
