---
title: "WHO"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{who}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
library(pddcs)
indicators <-
  indicatorlist[indicatorlist$source == 'who',]$indicator
```

This vignette provides details on the source data retrieved from WHO.

```{r, eval=TRUE, message=FALSE}
library(pddcs)
```

### About the source 

`fetch_indicator()` with `source = 'who'` retrieves data from the [Global Health Observatory](https://apps.who.int/gho/data/node.resources.api) provided by WHO. 

Note that some indicators are re-scaled before being returned by `fetch_indicator()`. Currently this applies to SH.MED.NUMW.P3, SH.MED.PHYS.ZS and SH.MED.BEDS.ZS, which are all re-scaled from values per 10,000 to values per 1,000. 

### Available indicators

The following `r length(indicators)` indicators are available for WHO.

```{r, echo=TRUE, message=FALSE}
library(dplyr)
data(indicatorlist)
indicatorlist %>% 
  filter(source == 'who') %>% 
  select(indicator)
```

### Examples

```{r, echo=TRUE, message=FALSE, eval=FALSE}
# Fetch data on hospital beds
df <- fetch_indicator('SH.MED.BEDS.ZS', source = 'who')

# Fetch data for mortality rate attributed to unsafe water
df <- fetch_indicator(
 indicator = c('SH.STA.WASH.P5', 'SH.STA.WASH.FE.P5', 'SH.STA.WASH.MA.P5'),
 source = 'who')
```

### References

WHO. Global Health Observatory. https://www.who.int/data/gho
